---
title: "Assignment 1"
author: "Manyi Luo - 1003799419"
date: "9/21/2019"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

#Solution

#1. Using words, give a title to summarise what your assignment #1 is about. Specify your explanatory variable and response variable and, give a brief explanation for your choice.

My assignment is about finding the relationship between height and head circumference using simple linear regression. So my title is: determining the relationship between height and head circumference. My explanatory variable (x) is height and the response variable (y) is head circumference. I choose height as explanatory variable (x) because in daily observations, taller people may have greater stature, so I am wondering whether higher people may have greater head circumference.

#2. Draw at most 3 plots to visually describe your data. Is your response variable approximately Normal?

```{r, echo = FALSE, fig.width=5.5, fig.height=3.5, message=FALSE}
#The result of doing a random sampling of 10 data from a1_60_data.csv (60 data in total) without replacement.
height = c(155,158,160,164,168,172,174,178,185,190)
headcircumference = c(55,58.3,56,57,56,57,58,58.5,59,62.5)

#The first plot
h_h_manyi = lm(headcircumference~height)
plot(height, headcircumference, xlab="Height in cm", ylab="Head circumference in cm",main="Scatterplot of height and head circumference 9419")
abline(h_h_manyi)

#The second plot
qqnorm(headcircumference, main="Q-Q plot of head circumference 9419")
qqline(headcircumference, main="Q-Q plot of head circumference 9419")

#The third plot
boxplot(headcircumference, main="boxplot of head circumference 9419")
```

Through the scatterplot, we can see that ten data points are located around the fitted line, eventhough some of them are above and some of them are below. There are two outliers identified, and the heights are 158cm and 190cm (head circumference are 58.3cm and 62.5cm) respectively. And based on the trend line, we can see height and head circumference follows a positive linear relationship: as height increases, head circumference increases. 

We can see the response variable (head circumference) is approximately normal based on the Q-Q plot. This is because, as data points evenly distributed through out the theoretical quantiles (x axis), most of them lies around the fitted line, and there is only one outstanding outlier located at theoretical quantile 1.5.

From the boxplot, we can clearly see that the distance between the 25th percentile and the median and the distance between the median and the 75th percentile are approximately the same, eventhough the distance between the median and the 75th percentile is slightly larger (since most of the data gather between 56-58cm), making it slightly left-skewed. There is only one outstanding outlier at 62cm, which doesn't really affect the distribution. Above all, the response variable is approximately Normal with a weak tendency of left-skewed.

#3. Numerically describe the centre, spread and any unusual points of your variables/data.

```{r, echo = FALSE}
summary(headcircumference)
IQR(headcircumference)
sd(headcircumference)
```

For variable headcircumference:

We can clearly see from the first chart that mean is 57.73cm and median is 57.5cm. And the interquartile range is calculated as 2.2.

Spread(standard deviation) of data is calculated as 2.104519.

```{r, echo = FALSE}
summary(height)
IQR(height)
sd(height)
```

For variable height:

We can clearly see from the first chart that mean is 170.4cm and median is 170cm. And the interquartile range is calculated as 16.

Spread(standard deviation) of data is calculated as 11.62564.

Also referring to the scatterplot from part (3), we can identify two unusual points of data, which are (158cm, 58.3cm) and (190cm, 62.5cm) respectively (height, headcircumference). We can clearly see that (190cm, 62.5cm) is an outlier because its the maximum, and it deviates a lot from the other data. Also, we might deduce that (158cm, 58.3cm) is an outlier from the data of variable headcircumference. We can see that the mean is slightly greater than the median, meaning that the existence of (158cm, 58.3cm) may drag the mean up. This also corresponds to our conclusion that the response variable is approximately Normal with a weak tendency of left-skewed. If data are sampled perfectly (eg: large enough sample size to dilute the influence of outliers), we might observe that the response variable becomes even closer to Normal.

#4. Fit and describe a simple linear regression model between head circumference and height.    
```{r, echo = FALSE}
h_h_manyi = lm(headcircumference~height)
summary(h_h_manyi)
```

We set up the linear regression model to describe the relationship between height and head circumference, as height is the explanatory variable and head circumference is the response variable. Therefore, our simple linear regression model can be expressed as $Y_i = \beta_0 + \beta_1 X_i + e_i$, where $Y_i$ is each individual's head circumference, $\beta_0$ is the true intercept of linear regression line, $\beta_1$ is the true slope of linear regression line, $X_i$ is each individual's height and last but not least, the $e_i$ is the random error associated with each data.

Using our ten sampling data points to find the fitted regression line, we can calculate the estimated $\beta_0$ (the intercept) is 32.58745, while the estimated $\beta_1$ equals 0.14755. Estimated $\beta_0$ represents that the head circumference is at least 32.58745cm (when height is 0cm, head circumference is 32.58745cm). The estimated $\beta_1$ represents the slope parameter, meaning that 1cm increase in height results 0.14755cm increase in head circumference. So the simple linear regression model can be represented as $\hat Y_i = 32.58745 + 0.14755 \hat x_i$.

#5. Are the regression parameters zero? Interpret the estimates of the regression parameters (p-value is held for four decimal places).

The p-value of estimated $\beta_0$ is 0.0009. This comes from setting $H_0$ as $\beta_0 = 0$ and $H_1$ as $\beta_0 ≠ 0$, and with a significance level of 5%, we have strong evidence to reject $H_0$ because 0.0009 is much more smaller than 0.05, meaning that we support $H_1$. $\beta_0$ is not likely to be zero.

The p-value of estimated $\beta_1$ is 0.0005. This is resulted from setting $H_0$ as $\beta_1 = 0$ and $H_1$ as $\beta_1 ≠ 0$, and with a significance level of 5%, we have strong evidence to reject $H_0$ because 0.0005 is much more smaller than 0.05, meaning that we support $H_1$. $\beta_1$ is not likely to be zero either.

#6. Visually and numerically describe the residuals of the fit in part (4) above. Use at most one plot (p-value is held for four decimal places.

```{r, echo = FALSE, fig.width=6, fig.height=4, message=FALSE}
h_h_manyi = lm(headcircumference~height)
plot(h_h_manyi, which = 1)
```

From the graph above, we can see that only 5 out of 10 data points lie around the trend line. We cannot identify a pattern in residuals' variation, as they all randomly spread out around the line. Therefore, the assumption of constant variance is satisfied.

And analyzing numeric data generated from part (4), the residual standard error is 1.293 on 8 degrees of freedom. The p-value of estimated residual standard error is 0.0041. This comes from setting $H_0$ as $residual\ error = 0$ and $H_1$ as $residual\ error  ≠ 0$, and with a significance level of 5%, we have evidence to reject $H_0$ because 0.0041 is smaller than 0.05, meaning that we support $H_1$. Residual error is not likely to be zero, which coincides with our graph. 

What's more, the multiple R-squared is 0.6644, meaning that 66.44% of change in head circumference can be explained by the change in height. This also means that the linear regression model we set up previously is well fitting.

#7. Identify a potential lurking variable. Describe at least one other issue or limitation of your fit.

A potential lurking variable is gender, which can influence both height and head circumference, distorting the linear relationship. For example, female might have lower average height and bigger head circumference comparing to male.

One of the limitation of my fit is that, because the sample size in this case is small (only 10 data randomly selected from 60 data), this might result rather high linearity or poor linearity. Statistically, if we want to get a representative result, we need to make our sample size greater or equal to 30.

#8. Identify another pair of variables to explore a simple linear regression model. Specify the response and explanatory variable.

Another pair of variables to explore might be weight and head circumference. But in order to make our sampling data more specific, we need to elimiate the effect of lurking variable, for exmaple: we need to specify sex before collecting data of weight and head circumference. In this case, the explanatory variable is weight and the response variable is head circumference. And the simple linear regression model can also be used to find the relationship between weight and head circumference, for example: as weight increases, head circumference also increases.

#Appendix

```
Q2
#The result of doing a random sampling of 10 data from a1_60_data.csv (60 data in total) without replacement.
height = c(155,158,160,164,168,172,174,178,185,190)
headcircumference = c(55,58.3,56,57,56,57,58,58.5,59,62.5)

#The first plot
h_h_manyi = lm(headcircumference~height)
plot(height, headcircumference, xlab="Height in cm", ylab="Head circumference in cm",main="Scatterplot of height and head circumference 9419")
abline(h_h_manyi)

#The second plot
qqnorm(headcircumference, main="Q-Q plot of head circumference 9419")
qqline(headcircumference, main="Q-Q plot of head circumference 9419")

#The third plot
boxplot(headcircumference, main="boxplot of head circumference 9419")

Q3
summary(headcircumference)
sd(headcircumference)
IQR(headcircumference)
summary(height)
sd(height)
IQR(height)

Q4
h_h_manyi = lm(headcircumference~height)
summary(h_h_manyi)

Q6
h_h_manyi = lm(headcircumference~height)
plot(h_h_manyi, which = 1)
```
